Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-08-09 extended_catalog/paper/extended_catalog.tex
markus changes

There will always be a small numerical discrepancy between the expected
photon distribution derived from a true point source and a very small
extended source due to numerical error in the convolution.  In most
situations, this error would be insignificant.  But in particular for
very bright sources, this numerical error has the potential to bias the
test statistic for the extension test.  the source with an extended shape
to the likelihood fitting it with its extension fixed to ${10^{-10}}\deg$.

@@

We found good agreement between the two methods. For the new extended
source candidates, table~\ref{alt_diff_model_results} will shows a
comparison of the test statistic \ts and \tsext computed using \pointlike
and \gtlike.  Unless explicitly mentioned, all \ts, \tsext, and spectral
parameters were calculated using \gtlike using \pointlike's positions
and extensions.

@@

Like extended sources described in section~\ref{gtlike_crosscheck},
the spectrum of the two point sources can be refit using \gtlike.
We quote the spectral values obtained from \gtlike using the best fit
positions found using \pointlike.  For the extended candidates found
by our search, we computed both \tsext and \tsinc.  The results will be
presented in table~\ref{dual_localization_results}.

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extension varying from $\sigma=0.1\deg$
to $2.0\deg$.  The LAT's threshold to extension is better for harder
sources with more high energy photons and the threshold is best for
sources with an extension $\approx0.5\deg$ and is significantly worse
for very small extended small.  Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only photons with energies between 1 \gev and 100 \gev.
Except for the hardest and largest sources, our detection threshold is
almost unchanged when including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in figure~\ref{four_plots_ic443}(b)
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\approx2.4$, almost the entire increase in \ts comes
from energies above 1 \gev.  On the other hand, other systematic errors
become increasingly important at low energy. For our extension search,
we use only only photons with energy above 1 \gev.

Figure~\ref{diff_factor_sensitivity} shows how the LAT's detection
threshold increases with increasing background. We compute the threshold
using 1, 10, and 100 times the Sreekumar-like isotropic background. The
latter factor approximately represents the intensity of the Galactic
diffuse emission near the Galactic center.  The figure shows the threshold
using only photons with energies between 1 \gev and 100 \gev and aldo
when using only photons with energies between 10 \gev and 100 \gev. We
are less sensitive to extended sources with higher background. Also,
given the same flux we are more sensitive to the source if those
photons have a higher energies.

@@

Finally, figure~\ref{all_sensitivity} shows the LAT's detection threshold
for all four spectral indices (1.5, 2, 2.5, and 3) against all three
background levels (1, 10, and 100 times the Sreekumar-like isotropic
background).  For a fixed 1 \gev to 100 \gev or 10 \gev to 100 \gev
flux, the LAT's detection threshold is only weakly depends upon the
spectral index.  This effect is most pronounced when using only photons
with energies above 10 \gev. The extension thresholds are tabulated in
table~\ref{all_sensitivity_table}.

@@

2FGL is a list of 1873 \gev sources detected by the LAT.  2FGL included
twelve previously published spatially extended sources but did not
attempt to fit the extension of these sources. Other than these sources,
all 2FGL sources were modeled as point sources and 2FGL did not attempt to
resolve the extension of new sources.  In this paper, we tested all 2FGL
sources for extension to search for new spatially extended \gev sources.

For this search, we build a model of the
sky consistent with 2FGL.   We used the same two
years of data from August 4, 2008 to August 1, 2010 and we used the same
Pass 7\_V6 (P7\_V6) Source class event selections and 
Instrument Response Functions (IRFs, \cite{lat_on_orbit_psf}).  We used the same Galactic diffuse,
isotropic diffuse, and earth limb emission model. The spectrum of the
Galactic and isotropic diffuse models were refit during the analysis
while the earth limb emission was fixed to it's predicted value.

As was shown, we gain little in sensitivity using photons with energies
below 1 \gev. On the other hand, the large PSF at low energy makes us
more susceptible to systematic errors arising from source confusion due
to multiple point sources and susceptible to incorrect modeling of the
Galactic diffuse emission. In addition, the Galactic diffuse emission
is more pronounced at lower energies due to its steep energy spectrum.
For that reason, we performed our search using photons only between 1
\gev and 100 \gev.  We used 4 energy bins per each logarithmic decade
in energy.

We also performed a search for extended sources using photons only
between 10 \gev and 100 \gev. Even though we are testing the same
sources, this approach is complimentary because the Galactic diffuse
emission is even less dominant above 10 \gev. Furthermore, source
confusion becomes less of a problem since most LAT sources are not
significant at energies above 10 \gev.  Using higher energies, we might
expect to detect harder sources in more complicated regions. This is
beneficial for regions near pulsars which cutoff before 10 \gev and was
successfully used in detecting HESS\,J1825-137 and MSH 15-52 with the LAT
(\cite{msh1552,fermi_hess_j1825}).

For each catalog source, we tested it for extension using \pointlike
assuming the source had a uniform radially symmetric disk spatial model.
We used a circular $10\deg$ region of interest centered on our source and
included all catalog sources within $15\deg$ of the source of interest.
We refit the spectral parameters of sources within $2\deg$ of the source.
We modeled each source with a power-law spectral model. This was reasonable because for a source to be significantly
extended, it would have to be fairly hard and fit reasonably well by
a power-law. Second, since we began the analysis at either 1 \gev or
10 \gev, we didn't expect there to be significant curvature over this
energy range.

Finally, when analyzing the region, we automatically removed other
2FGL sources which were within 0.5\deg of the source of interest. This was done because of
a concern that extended sources would end up 2FGL as
multiple point sources. These spurious sources would then
distort the extension fit.
For each extended candidate, we performed a dual localization procedure
(described in section~\ref{dual_localization_method} to test if the
source is extended. We select the sources where $\tsext>16$.

@@

Since extragalactic sources are typically too far away to be resolvable,
we expect most extended sources to be inside our Galaxy and most
likely on the Galactic plane.  Unfortunately, the \gev emission
in the Galactic plane is dominated by extremely structured Galactic
diffuse emission.  Finding sources on top of this emission is difficult
(\cite{first_diffuse_paper}) as has been discussed in 1FGL and 2FGL
(\cite{first_cat,second_cat}).  Furthermore, the Galactic plane is crowded
and it is often difficult to correctly model nearby sources.  Because of this,
finding a source with $\tsext>16$ is not a sufficient
criteria for resolving a source. For each extended source,
we perform several analysis crosschecks.

% 2FGL J1856.2+0450c = P72Y3047 for
As an example of why this is important, we describe the test of 
2FGL\,J1856.2+0450c for extension.  Figure~\ref{example_bad_fit} shows
a Galactic and isotropic diffuse emission subtracted
smoothed counts map of this source using photons with energy 
between 1 \gev and 100 \gev. 
In this region along the Galactic plane, there appears to be 
large-scale residual in the diffuse emission. As a result, 2FGL\,J1856.2+0450c 
fit to an extension of 1.83\deg and the result is statistically significant
with \tsext=45.4.  By looking at the residuals, it is clear that this
complicated region is not fit well even though the source's extension is
statistically significant. For this reason, we manually discard
the extension of this and similar sources.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

For each extended source candidate,
we took the spatial and spectral model found using \pointlike and
performed a crosscheck by refitting the spectral parameters using
\gtlike.  We used a square ROI $14.1\deg$ in diameter with $1/32\deg$
spatial bins and the same energy binning as \pointlike. Using \gtlike,
we found a second measure of \tsext, this time not fitting the extension
but using a fixed template with the extension determined by \pointlike.
We only considered a source to be significantly extended if \pointlike
and \gtlike both agreed.  For the new extended source candidates,
table~\ref{alt_diff_model_results} will show a comparison of of \tsext for
\gtlike and \pointlike.

For each candidate, we generated a map of residual \ts
by adding a new source of spectral index 2 into the region and finding
the increase in likelihood when fitting its flux. \ts maps are useful
to look for residual structure in the sky mode.  Figure~\ref{res_tsmaps}
shows a residual \ts map for the extended source IC443.

For each candidate, we generated plots of the sum of all counts
within a given distance of the source and compared it to the model
predictions of a point source.  The plots were binned uniformly
in $\Delta \theta^2$ so that an isotropic background would be flat.
An example radial integral plot is shown for the extended source IC443
in figure~\ref{four_plots_ic443}(c).  For each source, we also made
Galactic and extragalactic diffuse emission subtracted smoothed counts
maps. Figure~\ref{four_plots_ic443}(d) shows an example for IC443.

For each source, we tested to see if the region contained instead
multiple point sources by replacing the extended source with
two point sources and performed the dual localization procedure
described in section~\ref{dual_localization_method}.  We only
considered a source to be extended if $\tsext>\tsinc$.  For the new
extended sources, we will show the results of this dual localization in
table~\ref{dual_localization_results}.

Because of the high source density in the Galactic plane, for promising
candidates we often had to iteratively improve the model of our background
sources to obtain a better fit of the sources.  Often, several catalog
sources would fill in the emission of the extended source and would
have to be removed from the region.  Similarly, background sources
were often insignificant in the fit energy range and had to be removed.
The position of background sources would often have to be refit once the
extended source was fit.  When the model of the extended source coupled
strongly with nearby sources, we iteratively fit the extended source and
all nearby sources until the fit converged.  For each extended source,
we describe the modification that were required.

@@

We estimate a systematic error on the extension of a source due to
uncertainty in our knowledge of the angular resolution of the LAT.  Before
launch, the LAT PSF was determined by a detector simulation which was
verified in accelerator tests (\cite{atwood_LAT_mission}).  A discrepancy
was found in flight data above a few \gev in the PSF compared to the
shape of bright AGN.  A publication on this issue is in preparation
(\cite{lat_on_orbit_psf}).  Subsequently, the PSF was fit empirically to
bright AGN and this empirical parameterization is the default PSF used
in the P7\_V6 IRFs.  To account for this uncertainty in our knowledge of
the PSF, we refit our extended source candidates using the Monte Carlo
representation of the PSF and consider the difference in extension found
using the two PSFs as a systematic error on the extension of a source.
Because the Monte Carlo representation of the PSF is smaller, it leads
to unphysically higher \tsext values.  This same approach was used in
(\cite{ic443}).  At high energies, we believe that our parameterization
of the PSF from bright AGN is substantially better than the Monte Carlo
representation of the PSF so this systematic errors is conservative.

@@

We first present on our analysis of the twelve extended sources
included in 2FGL.
(\cite{second_cat}).
The six extended SNRs IC443, W28, W30, W44, W51C, and the
Cygnus Loop were included in 2FGL.  IC443, W28, W44, W51C were
first reported in (\cite{ic443,w28,w44,w51c}).  Detailed papers
by the LAT collaboration studying W30 and the Cygnus Loop are
still in preparation.  \gev emission from W30 was also studied in
(\cite{castro_and_slane_2010}).  Using photons with energies between
1 \gev and 100 \gev, our analysis significantly detected the extension
of all six SNRs. 
% 2 galaxies

Two nearby galaxies the Large Magalenic Cloud and the Small Magalenic
Cloud were included in 2FGL (\cite{lmc,smc}).  They were significantly
detected using photons with energies between 1 \gev and 100 \gev. Our
fit extension is comparable to the published result, but we note that
previous publication used a more complicated double Gaussian spatial
model when fitting the LMC.

@@

Nine extended sources not include in 2FGL were found using this search.
Three were found in our search using photons with energies between
1 \gev and 100 \gev and six were found using photons with energies between
10 \gev and 100 \gev.  The fit properties of these nine sources is summarized in
table~\ref{new_ext_srcs_table}.  This table includes for each source
the best fit position, extension, spectrum, source significance, and
significance of extension.

@@

This source is spatially coincident with the SNR RX J1713.7-3946 and was
recently reported by (\cite{rx_j1713_lat}).  To analyze this source,
we used the same background model as the recent LAT publication.
2FGL\,J1715.4-4024c is spatially coincident with Source A and was
moved to the published location of 
$(\text{RA},\text{Dec})=(258.84,-40.46)$. Source
B was added at 
$(\text{RA},\text{Dec})=(258.71,-38.70)$ and Source C was added at
$(\text{RA},\text{Dec})=(257.47,-39.75)$.  Figure \ref{2FGL_J1712.4-3941} shows
the position of these nearby sources on top of a contour of SNR RX
J1713.7-3946 seen in \tev by H.E.S.S (\cite{rx_j1713_hess}).

@@

This source is currently unidentifed. There are five nearby pulsars,
but they are not luminous enough to provide the required energy output
to power the \tev emission (\cite{closer_look_hess_j1614-518}).

@@

\todo[inline]{Finish the discussion.}

@@

Figure~\ref{compare_index_2FGL} shows a comparison of the spectral index
of LAT detected extended sources and of all sources in 2FGL. This, and
table~\label{known_extended_sources} and~\label{new_ext_srcs_table},
show that the LAT observes a population of hard (index $<2$)
extended sources above 10 \gev which connects to \tev sources.
Figure~\ref{hess_seds} shows the LAT energy spectrum of four of the
extended sources which were first discovered in the H.E.S.S. Galactic
plane survey (\cite{hess_plane_survey}) and have a matching size to LAT
extended sources.  The LAT energy spectrum connects smoothly to \tev.
This is also true of HESS\,J1825-137 (\ref{fermi_hess_j1825}) and RX
J1713.7-3946 (\cite{rx_j1713_lat}). We conclude that the \gev and \tev
emission from these sources is coming from the same emission mechanism.

@@

Figure~\ref{gev_vs_tev_histogram} compares the size of the 21 extended
LAT sources to the 42 extended H.E.S.S. sources.  Because of the large
field of view and all sky coverage, the LAT can more easily measure
larger sources including the LMC and SMC.  On the other hand, the
smaller angular resolution of Air Cherenkov detectors allows them to
measure a population of extended sources below our sensitivity limit
(currently at about $\approx0.2\deg$).  Fermi has a five year nominal
livetime with a goal of ten years of operation. With increased exposure,
the LAT will likely begin to resolving these smaller \tev sources.

Furthermore, our sensitivity to these extended sources will likely improve
faster than $\sqrt{t}$ because at the highest energies, our background is
relatively small and we are in the Poisson instead of Gaussian limit.
Future improvements in the diffuse emission model will improve our
sensitivity by allowing this analysis to go to lower energies.

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/four_plots_ic443.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/four_plots_ic443/v1/run.py
    \caption{
    All plots are for the SNR IC443 using two years of data.
    % ts vs extension
    Plot (a) is the change in \ts when varying the
    extension of the source.  
    % ts_ext vs energy
    Plot (b) is \tsext in twelve energy bins. This plot shows that there is
    little gain in sensitivity to the extension using photons
    with energies below 1 \gev, even for a source with an index $\approx2.4$.
    % radial integral
    Plot (c) is the sum of all counts within a given radial distance
    from the source binned uniformly in $\theta^2$. 
    Overlaid on the plot is the model predicted counts assuming that
    the source is both an extended source and a point source
    % smoothed counts
    Plot (d) is a counts map of the source with the Galactic and
    isotropic diffuse emission subtracted and smoothed by a $0.1\deg$
    Gaussian kernel.  The inset is the predicted emission from a point
    source of the same spectrum and intensity.  The PSF is smoothed
    by the same kernel.  Plots (a), (c), and (d) use only photons with
    energies between 1 \gev and 100 \gev. Plots (c) and (d)
    use only photons which converted in the front
    part of the detector and have an improved angular resolution.
    }
    \label{four_plots_ic443}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/compare_disk_gauss.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_disk_gauss/v2/compare_disk_gauss.pdf
    \end{center}
    \caption{
    A comparison of the Gaussian and uniform disk PDFs before and
    after convolving with the PSF.  The black line is the PSF that
    would be observed for a power-law source of spectral index 2. The
    dashed red line is the probability density of a Gaussian with
    $\rsixeight=0.5\deg$, which is just a straight line on this plot.
    The dashed blue line is the probability density for a uniform disk
    with $\rsixeight=0.5\deg$.  Finally, the solid red and blue lines
    show the convolution of the Gaussian and uniform disk with the PSF.
    Plot (a) shows the analysis in the energy range from 1 \gev to 100
    \gev while plot (b) shows the analysis only in the energy range
    from 10 \gev to 100 \gev. This plot demonstrates that different
    spatial models when smoothed by the PSF look similar.  The $y$
    axis is the log of the intensity and the $x$ axis is $\theta^2$.
    This plot is discussed further in section~\ref{compare_source_size}.
    }\label{compare_disk_gauss}
  \end{figure}

@@

  \caption{An example residual test statistic map generated for the
  supernova remnant IC443 using photons with energy between 1 \gev
  and 100 \gev. Plot (a) is the residual \ts fitting IC443 as a point
  source and plot (b) is the residual TS fitting IC443 as an extended
  source. The crosses in the plot represent all of the sources modeled
  in the region. Since there is residual TS greater than 160 fitting
  with the point hypothesis so it is not a good model of the source.
  On the other hand, the residual fitting IC443 as an extended source
  is much lower because IC443 is better fit by an extended source. These
  plots were generated for all extended source candidates to help validate
  the analysis.}
  \label{res_tsmaps}
  \end{center}
\end{figure}

@@

      list are the extended sources analyzed using photons with energies
      between 1 \gev and 100 \gev.  The lower table are the sources
      analyzed using photons with energies between 10 \gev and 100 \gev.

>>>>

There will always be a small numerical discrepancy between the expected
photon distribution derived from a true point source and a very small
extended source due to numerical error in the convolution.  In most
situations, this error would be insignificant.  But in particular for
very bright sources, this numerical error has the potential to bias the
test statistic for the extension test. The source with an extended shape
to the likelihood fitting it with its extension fixed to ${10^{-10}}\deg$.

@@

We found good agreement between the two methods. For the new extended
source candidates, table~\ref{alt_diff_model_results} shows a
comparison of the test statistic \ts and \tsext computed using \pointlike
and \gtlike.  Unless explicitly mentioned, all \ts, \tsext, and spectral
parameters were calculated using \gtlike using \pointlike's positions
and extensions.

@@

Like extended sources described in section~\ref{gtlike_crosscheck},
the spectrum of the two point sources can be refit using \gtlike.
We quote the spectral values obtained from \gtlike using the best fit
positions found using \pointlike.  For the extended candidates found
by our search, we computed both \tsext and \tsinc.  The results are
presented in table~\ref{dual_localization_results}.

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extension varying from $\sigma=0.1\deg$
to $2.0\deg$.  The LAT's threshold for a significant detection
of an extension of a LAT souce drops quickly with
increasing source size and reaches a minimum around 0.5\deg. 
Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only 1-100\gev photons.
Except for the hardest and largest sources, our detection threshold is
almost unchanged when including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in figure~\ref{four_plots_ic443}(b)
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\approx2.4$, almost the entire increase in \ts comes
from energies above 1 \gev.  On the other hand, other systematic errors
become increasingly important at low energy. For our extension search,
we use therefore only photons with energies above 1 \gev.

Figure~\ref{diff_factor_sensitivity} shows how the LAT's detection
threshold increases with increasing background. We compute the threshold
using 1, 10, and 100 times the Sreekumar-like isotropic background. The
latter factor approximately represents the intensity of the Galactic
diffuse emission near the Galactic center.  The figure shows the threshold
using 1-100 \gev photons and also
using only 10-100 \gev photons. We
are less sensitive to extended sources with higher background. Also,
given the same flux we are more sensitive to the source if those
photons have a higher energies.

@@

Figure 7 shows the flux threshold as a function of source extension for
different background levels (1x, 10x, and 100x the
nominal background), different spectral indices, and two
different energy bands (1-100 \gev and 10-100 \gev).
For a fixed 1-100 \gev or 10-100 \gev
flux, the LAT's detection threshold is only weakly depends upon the
spectral index.  This effect is most pronounced when using only photons
with energies above 10 \gev. The extension thresholds are tabulated in
table~\ref{all_sensitivity_table}.

@@

We test all sources in the 2FGL catalog for spatial extendedness.
2FGL is a catalog of 1873 sources detected by the LAT.  It included
twelve previously published spatially extended sources but did not
attempt to fit the extension of these sources. Other than these sources,
all 2FGL sources were modeled as point sources and 2FGL did not attempt
to resolve the extension of new sources.

Our analysis setup is closely related to the setup used for the
derivation of the catalog We used the same two years of data from August
4, 2008 to August 1, 2010 and we used the same Pass 7\_V6 (P7\_V6)
Source class event selection and Instrument Response Functions (IRFs,
\cite{lat_on_orbit_psf}).  We used the same models to describe the
background from Galactic diffuse, isotropic, and earth limb emission.
The galactic diffuse emission was scaled by a power-law and the isotropic
component's normalization was left free.

As was shown, we gain little in sensitivity using photons with energies
below 1 \gev. On the other hand, the large PSF at low energy makes us
more susceptible to systematic errors arising from source confusion due
to multiple point sources and susceptible to incorrect modeling of the
Galactic diffuse emission. In addition, the Galactic diffuse emission
is more pronounced at lower energies due to its steep energy spectrum.
For that reason, we performed our search using photons only between 1
\gev and 100 \gev.

We also performed a search for extended sources using only 10-100 \gev photons. 
Even though we are testing the same
sources, this approach is complimentary because the Galactic diffuse
emission is even less dominant above 10 \gev. Furthermore, source
confusion becomes less of a problem since most LAT sources are not
significant at energies above 10 \gev.  Using higher energies, we might
expect to detect harder sources in more complicated regions. This is
beneficial for regions near pulsars which cutoff before 10 \gev and was
successfully used in detecting HESS\,J1825-137 and MSH 15-52 with the LAT
(\cite{msh1552,fermi_hess_j1825}).

We tested each source for extension using
\pointlike
assuming the source had a uniform radially symmetric surface brightness.
We used a circular $10\deg$ region of interest centered on our source and
included all catalog sources within $15\deg$ of the source of interest
in our background model.
We refit the spectral parameters of sources within $2\deg$ of the source
to avoid potential biases of the extension parameters of the source of
interest due to close-by background sources.
We modeled each source with a power-law spectral model.

Finally, when analyzing the region, we automatically removed other 2FGL
sources which were within 0.5\deg of the source of interest. This was
done to avoid a concern that extended sources would end up in 2FGL as
multiple point sources. These spurious sources would then distort the
extension fit.  Instead, when a source is found to be significantly
extended (i.e. $\tsext>16$), we perform the dual localization procedure
(described in section~\ref{dual_localization_method} to compare the
extended source hypothesis to the hypothesis of two independent point
sources. Only sources with $\tsext>\tsinc$ are considered as extended.

@@

Based on size vs. distance considerations,
we expect most extended sources to be inside our Galaxy.  Unfortunately, the \gev emission
in the Galactic plane is dominated by extremely structured 
diffuse emission from the interactions of cosmic rays with the
intergalactic medium.  Finding sources on top of this emission is difficult
(\cite{first_diffuse_paper}) as has been discussed in 1FGL and 2FGL
(\cite{first_cat,second_cat}).  Furthermore, the Galactic plane is crowded
and it is often difficult to correctly model nearby sources.  Because of this,
finding a source with $\tsext>16$ is not a sufficient
criteria for resolving a source. For each extended source,
we perform several analysis crosschecks.

For each candidate, we generated a map of residual \ts by adding a new
source of spectral index 2 into the region and finding the increase in
likelihood when fitting its flux. \ts maps are useful to look for residual
structure in the sky mode.  Figure~\ref{res_tsmaps} shows a residual \ts
map for the extended source IC443.  We also generated plots of the sum of
all counts within a given distance of the source and compared it to the
model predictions of a point source.  The plots were binned uniformly
in $\Delta \theta^2$ so that an isotropic background would be flat.
An example radial integral plot is shown for the extended source IC443
in figure~\ref{four_plots_ic443}(c).  For each source, we also made
Galactic and extragalactic diffuse emission subtracted smoothed counts
maps. Figure~\ref{four_plots_ic443}(d) shows an example for IC443.

% 2FGL J1856.2+0450c = P72Y3047 for
We then inspected the residual \ts, smoothed counts maps, and radial
integral distributions to identify cases in which the extension fit
is clearly biased by large-scale residuals in the diffuse emission
and hence the extension measurement unreliable.  An example of such a
case is shown in figure~\ref{example_bad_fit}. It shows a Galactic and
isotropic diffuse emission subtracted smoothed counts map of this source
using 1-100 \gev photons.  In this region along the Galactic plane,
there appears to be large-scale residual in the diffuse emission. As
a result, 2FGL\,J1856.2+0450c fit to an extension of 1.83\deg and the
result is statistically significant with \tsext=45.4.  By looking at
the residuals, it is clear that this complicated region is not fit well
even though the source's extension is statistically significant. We do
not report sources as extended that fail this inspection.
% information came from
% /nfs/slac/g/ki/ki03/lande/extended_catalog/2FGL/v15/standard_analysis/spectral_emin_1000_v1/P72Y3047/v1/results_followup_P72Y3047.yaml

For the remaining candidates, 
we took the spatial and spectral model found using \pointlike and
redetermined the spectral 
parameters using
\gtlike.  We used 
the `binned likelihood' mode of \gtlike on a $14\deg\times14\deg$
ROI with a pixel size of 0.03\deg.
We obtained a second measure of \tsext from \gtlike for the extension
and position of the source as determined by \pointlike.
We only considered a source to be significantly extended if $\tsext>16$ with
both \pointlike and \gtlike.  For the new extended source candidates,
table~\ref{alt_diff_model_results} shows a comparison of of \tsext for
\gtlike and \pointlike.

Because of the high source density in the Galactic plane, for promising
candidates we often had to iteratively improve the model of our background
sources to obtain a better fit of the sources.  Several catalog
sources would often fill in the emission of the extended source and would
have to be removed from the region.  Similarly, background sources
were often insignificant in the fit energy range and had to be removed.
The position of background sources would often have to be refit once the
extended source was fit.  When the model of the extended source coupled
strongly with nearby sources, we iteratively fit the extended source and
all nearby sources until the fit converged.  For each extended source,
we describe the modifications of the background model compared to the model
used in the creation of the 2FGL catalog that were required.

@@

We estimate a systematic error on the extension of a source due to
uncertainty in our knowledge of the angular resolution of the LAT.  Before
launch, the LAT PSF was determined by a detector simulation which was
verified in accelerator tests (\cite{atwood_LAT_mission}).  A discrepancy
was found in flight data above a few \gev in the PSF compared to the
shape of bright AGN.  A publication on this issue is in preparation
(\cite{lat_on_orbit_psf}).  Subsequently, the PSF was fit empirically to
bright AGN and this empirical parameterization is the default PSF used
in the P7\_V6 IRFs.  To account for this uncertainty in our knowledge of
the PSF, we refit our extended source candidates using the Monte Carlo
representation of the PSF and consider the difference in extension found
using the two PSFs as a systematic error on the extension of a source.
The same approach was used in
(\cite{ic443}).  At high energies, we believe that our parameterization
of the PSF from bright AGN is substantially better than the Monte Carlo
representation of the PSF so this systematic errors is conservative.

@@

We first present on our analysis of the twelve extended sources
included in 2FGL.
(\cite{second_cat}).
The six extended SNRs IC443, W28, W30, W44, W51C, and the
Cygnus Loop were included in 2FGL.  IC443, W28, W44, W51C were
first reported in (\cite{ic443,w28,w44,w51c}).  Detailed papers by the LAT
collaboration studying W30 and the Cygnus Loop are still in preparation.
\gev emission from W30 was also studied in (\cite{castro_and_slane_2010}).
Using 1-100 \gev photons, our analysis significantly detected the
extension of all six SNRs.
% 2 galaxies

Two nearby galaxies the Large Magalenic Cloud and the Small Magalenic
Cloud were included in 2FGL (\cite{lmc,smc}).  They were significantly
detected using 1-100 \gev photons. Our
fit extension is comparable to the published result, but we note that
previous publication used a more complicated double Gaussian spatial
model when fitting the LMC.

@@

Nine extended sources not include in 2FGL were found using this search.
Three were found in our search using 1-100 \gev photons and 
six were found in our search using 10-100 \gev photons.  
The fit properties of these nine sources is summarized in
table~\ref{new_ext_srcs_table}.  This table includes for each source
the best fit position, extension, spectrum, source significance, and
significance of extension.

@@

This source is spatially coincident with the SNR RX J1713.7-3946
and was recently reported by (\cite{rx_j1713_lat}).  Figure
\ref{2FGL_J1712.4-3941} overlays H.E.S.S. \tev contours of SNR RX
J1713.7-3946 from (\cite{rx_j1713_hess}).  To analyze this source,
we used the same background model as the recent LAT publication.
2FGL\,J1715.4-4024c is spatially coincident with Source A and was
moved to $(\text{RA},\text{Dec})=(258.84,-40.46)$. Source B was added
at $(\text{RA},\text{Dec})=(258.71,-38.70)$ and Source C was added at
$(\text{RA},\text{Dec})=(257.47,-39.75)$.

@@

This source is currently unidentifed. There are five nearby pulsars,
but note are luminous enough to provide the required energy output
to power the \tev emission (\cite{closer_look_hess_j1614-518}).

HESS J1614-518 is spatially coincident with a young open cluster Pismis
22 (\cite{hess_1614_landi_atel,closer_look_hess_j1614-518}).
\gev and \tev emission could arrise from
hadronic acceleration in the winds of massive stars
\cite{cosmic_rays_and_gamma-rays_from_ob_stars,gamma-ray_production_in_young_open_clusters}.
Pisim 22 is at a distance $\approx1$ kpc and can energetically account
for the \tev emission (\cite{closer_look_hess_j1614-518}), but
it is difficult to explain the difference in size and positional offset
between Pismis 22 and HESS\,J1614-518.

@@

@@

Figure~\ref{compare_index_2FGL} shows a comparison of the spectral index
of LAT detected extended sources and of all sources in 2FGL. This, and
table~\ref{known_extended_sources} and~\ref{new_ext_srcs_table},
show that the LAT observes a population of hard (index $<2$)
extended sources above 10 \gev which connects to \tev sources.
Figure~\ref{hess_seds} shows the LAT energy spectrum of four of the
extended sources which were first discovered in the H.E.S.S. Galactic
plane survey (\cite{hess_plane_survey}) and have a matching size to LAT
extended sources.  The LAT energy spectrum connects smoothly to \tev.
This is also true of HESS\,J1825-137 (\ref{fermi_hess_j1825}) and RX
J1713.7-3946 (\cite{rx_j1713_lat}). We conclude that the \gev and \tev
emission from these sources is coming from the same emission mechanism.

@@

Figure~\ref{gev_vs_tev_histogram} compares the size of the 21 extended
LAT sources to the 42 extended H.E.S.S. sources.  Because of the large
field of view and all sky coverage, the LAT can more easily measure
larger sources including the LMC and SMC.  On the other hand, the
smaller angular resolution of Air Cherenkov detectors allows them to
measure a population of extended sources below our sensitivity limit
(currently at about $\approx0.2\deg$).  Fermi has a five year nominal
livetime with a goal of ten years of operation. With increased exposure,
the LAT will likely begin to resolving these smaller \tev sources.
As figure~\ref{time_sensitivity} showed, our sensitivity to these smaller
sources will actually improve faster than $\sqrt{t}$.

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/four_plots_ic443.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/four_plots_ic443/v1/run.py
    \caption{
    All plots are for the SNR IC443 using two years of data.
    % ts vs extension
    Plot (a) is the change in \ts when varying the
    extension of the source.  
    % ts_ext vs energy
    Plot (b) is \tsext in twelve energy bins. This plot shows that there is
    little gain in sensitivity to the extension using photons
    with energies below 1 \gev, even for a source with an index $\approx2.4$.
    % radial integral
    Plot (c) is the sum of all counts within a given radial distance
    from the source binned uniformly in $\theta^2$. 
    Overlaid on the plot is the model predicted counts assuming that
    the source is both an extended source and a point source
    % smoothed counts
    Plot (d) is a counts map of the source with the Galactic and
    isotropic diffuse emission subtracted and smoothed by a $0.1\deg$
    Gaussian kernel.  The inset is the predicted emission from a point
    source of the same spectrum and intensity.  The PSF is smoothed
    by the same kernel.  Plots (a), (c), and (d) use only 1-100 \gev photons. 
    Plots (c) and (d)
    use only photons which converted in the front
    part of the detector and have an improved angular resolution.
    }
    \label{four_plots_ic443}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/compare_disk_gauss.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_disk_gauss/v2/compare_disk_gauss.pdf
    \end{center}
    \caption{
    A comparison of the Gaussian and uniform disk PDFs before and
    after convolving with the PSF.  The black line is the PSF that
    would be observed for a power-law source of spectral index 2. The
    dashed red line is the probability density of a Gaussian with
    $\rsixeight=0.5\deg$, which is just a straight line on this plot.
    The dashed blue line is the probability density for a uniform disk
    with $\rsixeight=0.5\deg$.  Finally, the solid red and blue lines
    show the convolution of the Gaussian and uniform disk with the PSF.
    Plot (a) uses a 1-100 \gev energy range while plot (b)
    uses a 10-100 \gev enregy range. This plot demonstrates that different
    spatial models when smoothed by the PSF look similar.  The $y$
    axis is the log of the intensity and the $x$ axis is $\theta^2$.
    This plot is discussed further in section~\ref{compare_source_size}.
    }\label{compare_disk_gauss}
  \end{figure}

@@

  \caption{An example residual test statistic map generated for the
  supernova remnant IC443 using 1-100 \gev photons. 
  Plot (a) is the residual \ts fitting IC443 as a point
  source and plot (b) is the residual TS fitting IC443 as an extended
  source. The crosses in the plot represent all of the sources modeled
  in the region. Since there is residual TS greater than 160 fitting
  with the point hypothesis so it is not a good model of the source.
  On the other hand, the residual fitting IC443 as an extended source
  is much lower because IC443 is better fit by an extended source. These
  plots were generated for all extended source candidates to help validate
  the analysis.}
  \label{res_tsmaps}
  \end{center}
\end{figure}

@@

      list are the extended sources analyzed using 1-100 \gev photons.  
      The lower table are the sources
      analyzed using 10-100 \gev photons.
